PyBOP: A Python package for battery model optimisation and parameterisation

### 2024-11-29 joss/paper.preprint.tex
Adds Agriya w/ correct affiliation, updts tex pre-print

  pdfauthor={Brady Planden, Nicola E. Courtier, Martin Robinson, Ferran
Brosa Planella, David A. Howey},

@@

\usepackage[affil-it]{authblk}
\usepackage{orcidlink}
%% \renewcommand\Authsep{, }
\setlength{\affilsep}{1em}
\author[1%
  %
  \ensuremath\mathparagraph]{Brady Planden%
    \,\orcidlink{0000-0002-1082-9125}\,%
    }
\author[1,2%
  %
  ]{Nicola E. Courtier%
    \,\orcidlink{0000-0002-5714-1096}\,%
    }
\author[3%
  %
  ]{Martin Robinson%
    \,\orcidlink{0000-0002-1572-6782}\,%
    }
\author[2,4%
  %
  ]{Ferran Brosa Planella%
    \,\orcidlink{0000-0001-6363-2812}\,%
    }
\author[1,2%
  %
  ]{David A. Howey%
    \,\orcidlink{0000-0002-0620-3955}\,%
    }

\affil[1]{Department of Engineering Science, University of Oxford,
Oxford, UK%
  }
\affil[2]{The Faraday Institution, Harwell Campus, Didcot, UK%
  }
\affil[3]{Research Software Engineering Group, University of Oxford,
Oxford, UK%
  }
\affil[4]{Mathematics Institute, University of Warwick, Coventry, UK%
  }
\affil[$\mathparagraph$]{Corresponding author: %
   %
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\date{28 November 2024}

@@

\texttt{PyBOP} is a Python package designed to provide a user-friendly,
object-oriented interface for optimising battery models. \texttt{PyBOP}
leverages the open-source \texttt{PyBaMM} package
(\citeproc{ref-Sulzer:2021}{Sulzer et al., 2021}) to formulate and solve
of these battery models. \texttt{PyBOP} is intended to serve a broad
audience of students, engineers, and researchers in both academia and
the battery industry by enabling the use of predictive battery models
where previously this was not possible. \texttt{PyBOP} emphasises clear
and informative diagnostics and workflows for users of varying
expertise, by providing advanced optimisation and sampling algorithms.
These methods are provided through interfaces to \texttt{PINTS}
(\citeproc{ref-Clerx:2019}{Clerx et al., 2019}), \texttt{SciPy}
(\citeproc{ref-SciPy:2020}{Virtanen et al., 2020}), in addition to the
PyBOP's own algorithms such as Adaptive Moment Estimation with Weight
Decay (AdamW), and Cuckoo search.

@@

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/sim-landscape.png}
\caption{The cost landscape for the parameterisation problem with a root
mean squared error cost function. \label{fig:inference-time-landscape}}
\end{figure}

As gradient information is available for this problem, the choice of
distance-based cost function and optimiser is not constrained. Due to
the different magnitudes of the two parameters, we apply the logarithmic
parameter transformation offered by \texttt{PyBOP}. This transforms the
optimisers search space of the optimiser to allow for a common step size
between the parameters, which is generally is not required, but improves
convergence in this problem. As a demonstration of the parameterisation
capabilities of \texttt{PyBOP}, \autoref{fig:convergence-min-max} (left)
shows the rate of convergence for each of the distance-minimising cost
functions, while \autoref{fig:convergence-min-max} (right) shows
analogous results for maximising a likelihood. The optimisation is
performed with SciPy Minimize using the gradient-based L-BFGS-B method.

@@

>>>>

  pdfauthor={Brady Planden, Nicola E. Courtier, Martin Robinson, Agriya
Khetarpal, Ferran Brosa Planella, David A. Howey},

@@

\usepackage[affil-it]{authblk}
\usepackage{orcidlink}
%% \renewcommand\Authsep{, }
\setlength{\affilsep}{1em}
\author[1%
  %
  \ensuremath\mathparagraph]{Brady Planden%
    \,\orcidlink{0000-0002-1082-9125}\,%
    }
\author[1,2%
  %
  ]{Nicola E. Courtier%
    \,\orcidlink{0000-0002-5714-1096}\,%
    }
\author[3%
  %
  ]{Martin Robinson%
    \,\orcidlink{0000-0002-1572-6782}\,%
    }
\author[4%
  %
  ]{Agriya Khetarpal%
    \,\orcidlink{0000-0002-1112-1786}\,%
    }
\author[2,5%
  %
  ]{Ferran Brosa Planella%
    \,\orcidlink{0000-0001-6363-2812}\,%
    }
\author[1,2%
  %
  ]{David A. Howey%
    \,\orcidlink{0000-0002-0620-3955}\,%
    }

\affil[1]{Department of Engineering Science, University of Oxford,
Oxford, UK%
  }
\affil[2]{The Faraday Institution, Harwell Campus, Didcot, UK%
  }
\affil[3]{Research Software Engineering Group, University of Oxford,
Oxford, UK%
  }
\affil[4]{Quansight Labs%
  }
\affil[5]{Mathematics Institute, University of Warwick, Coventry, UK%
  }
\affil[$\mathparagraph$]{Corresponding author: %
   %
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\date{29 November 2024}

@@

\texttt{PyBOP} is a Python package designed to provide a user-friendly,
object-oriented interface for optimising battery models. \texttt{PyBOP}
leverages the open-source \texttt{PyBaMM} package
(\citeproc{ref-Sulzer:2021}{Sulzer et al., 2021}) to formulate and solve
of these battery models. \texttt{PyBOP} is intended to serve a broad
audience of students, engineers, and researchers in both academia and
the battery industry by enabling the use of predictive battery models
where previously this was not possible. \texttt{PyBOP} emphasises clear
and informative diagnostics and workflows for users of varying
expertise, by providing advanced optimisation and sampling algorithms.
These methods are provided through interfaces to \texttt{PINTS}
(\citeproc{ref-Clerx:2019}{Clerx et al., 2019}), \texttt{SciPy}
(\citeproc{ref-SciPy:2020}{Virtanen et al., 2020}), in addition to the
PyBOP's own algorithms such as Adaptive Moment Estimation with Weight
Decay (AdamW), Gradient descent, and Cuckoo search.

@@

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/sim-landscape.png}
\caption{The cost landscape for the time-series parameterisation problem
with a root mean squared error cost function.
\label{fig:inference-time-landscape}}
\end{figure}

As mentioned above, \texttt{PyBOP} also provides inference and
optimisation capabilities through numerical impedance spectroscopy. This
is based on the methods presented in the \texttt{pybamm-eis} package and
allows fast impedance computation by inversion of the sparse PyBaMM mass
matrix after scaling and subtraction of the auto-differentiated Jacobian
matrix. More information about this procedure is available in
(\citeproc{ref-pybamm-eis}{Dhoot et al., n.d.}). The
\autoref{fig:impedance-landscape} below shows the numerical impedance
prediction available in \texttt{PyBOP} alongside the cost landscape
constructed for the inference task. At the time of publication,
gradient-based optimisation and sampling methods are not available when
using an impedance workflow.

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/impedance.png}
\caption{The cost landscape for the frequency domain impedance
parameterisation problem with a root mean squared error cost function at
5\% SOC. \label{fig:impedance-landscape}}
\end{figure}

Moving forward, we will continue with identification in the time-domain.
As gradient information is available for this problem, the choice of
distance-based cost function and optimiser is not constrained. Due to
the different magnitudes of the two parameters, we apply the logarithmic
parameter transformation offered by \texttt{PyBOP}. This transforms the
optimisers search space of the optimiser to allow for a common step size
between the parameters, which is generally is not required, but improves
convergence in this problem. As a demonstration of the parameterisation
capabilities of \texttt{PyBOP}, \autoref{fig:convergence-min-max} (left)
shows the rate of convergence for each of the distance-minimising cost
functions, while \autoref{fig:convergence-min-max} (right) shows
analogous results for maximising a likelihood. The optimisation is
performed with SciPy Minimize using the gradient-based L-BFGS-B method.

@@

\bibitem[\citeproctext]{ref-pybamm-eis}
Dhoot, R., Timms, R., \& Please, C. (n.d.). \emph{PyBaMM EIS: Efficient
linear algebra methods to determine li-ion battery behaviour} (Version
0.1.4). \url{https://www.github.com/pybamm-team/pybamm-eis}
